---
title: "Data 608 Module 1"
author: Aaron Zalki
date: 09/06/2020
output:
  html_document:
    highlight: pygments
    theme: journal
    toc: true
    toc_float:
      collapsed: true
---

```{r, warnings=FALSE, message=FALSE}
library(corrgram)
library(ggplot2)
library(tidyr)
library(dplyr)

```

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}
#github path
inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```

```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

Correlation charts from the `corrgram library` help show if the variables are related to one another. It does not seem that they have a simple linear relationship. There seems to be a relationship between Employees and Revenue, which we can look at below.



```{r}
# Insert your code here, create more chunks as necessary

corrgram(inc, order=TRUE, lower.panel=panel.ellipse,
  upper.panel=panel.pts, text.panel=panel.txt,
  diag.panel=panel.minmax)
```

## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r}

# Answer Question 1 here
state = inc %>%
  #group data by state
  group_by(State) %>%
  #provide counts
  count(State)%>%
  #sort from highest to lowest
  arrange(desc(n))
#top 6 states
head(state)

```

```{r, fig.height = 12, fig.width = 10}

states_plot <- ggplot(state, aes(x=reorder(State, n), y=n, fill=n))
states_plot + geom_bar(stat="identity", width=0.4, position = position_dodge(width=0.5)) + 
coord_flip() + 
#label axis
labs(x = "State", y = "Number of Companies")


```

## Question 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r}

# Answer Question 2 here
#full data
inc <- inc[complete.cases(inc),]
#filter for New York, since this is the third highest
ny = inc %>%
  filter(State == "NY")

#plot data to show the average employment by industry for companies in New York
ny_plot <- ggplot(ny, aes(reorder(Industry,Employees,mean), Employees))
ny_plot <- ny_plot + geom_boxplot() + coord_flip() + 
  labs(x = "Industry", y = "Number of Employees") 
ny_plot
```


```{r}
#adjust plot
ny_plot + scale_y_log10()

```

### Remove Outliers


```{r}
gaussian_plot <- ggplot(ny, aes(Employees))
gaussian_plot + geom_density(kernel = "gaussian") 
```


```{r}
subs3t_ny <- select (ny, Industry, Employees)
head(subs3t_ny %>% arrange(desc(Employees)))

```
**Let's filter the data to only show employees less than or equal to 3000. We can see above there are high values (10000 and 32000)**

```{r}

ny_filtered = ny %>%
  #our filter criteria
  filter(Employees <= 3000)

#plot data without outliers

ny_plot2 <- ggplot(ny_filtered, aes(reorder(Industry,Employees,mean), Employees))
ny_plot2 <- ny_plot2 + geom_boxplot() + coord_flip() + 
  labs(x = "Industry", y = "Number of Employees") 
ny_plot2
```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r, message = FALSE, warnings = FALSE}

# Answer Question 3 here

#full data
inc <- inc[complete.cases(inc),]

industry = inc %>%
  #group data by industry
  group_by(Industry) %>%
  #aggregate revenue and employee data
  summarise(Revenue=sum(Revenue), Employees=sum(Employees)) %>%
  #divide values to find revenue per employee
  mutate(per_employee = Revenue/Employees)


#plot data to show which industries generate the most revenue per employee

revenue_plot <- ggplot(industry, aes(x=reorder(Industry, per_employee), 
                                     y=per_employee, fill=per_employee))
revenue_plot + geom_bar(stat="identity") + coord_flip() + 
  labs(x = "Industry", y = "Revenue per Employee")

```


```{r}

revenue_data = inc %>%
  mutate(per_employee = Revenue/Employees)

mean_plot <- ggplot(revenue_data, aes(reorder(Industry,per_employee,mean), per_employee))
mean_plot <- mean_plot + geom_boxplot() + coord_flip() +   
  labs(x = "Industry", y = "Revenue per Employee") 
mean_plot
```

```{r}
#adjust plot
mean_plot + scale_y_log10()

```
